---
title: "R course Peer Assessment 1"
output: html_document
date: "Wednesday, April 15, 2015"
---

### Loading and preprocessing the data ###

setwd to locate the activity.csv, using the read.csv with header to load the data.

```{r echo = TRUE}
Sys.setlocale("LC_TIME", "C")
library(dplyr)
library(ggplot2)
setwd("F:/Rprog/repdata-data-activity")
file_url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

if (!file.exists("activity.zip")){
        download.file(file_url, destfile = "activity.zip")
       unzip("activity.zip")
}

activity <- read.csv("activity.csv", header = TRUE)
```

### What is mean total number of steps taken per day? ###

using the sumarize of dplyr to sum the total steps grouped by date

```{r echo = TRUE}
### plot the histogram
Hist <- function(dataset){      
hist(dataset$total.steps,
     xlab = "Steps Per Day",
     main = "Histogram of 'Steps Per Day'"
     )

}

### Calculate the total number of steps taken per day
Bydate <- function(dataset){
        summarize(group_by(dataset, date), sum(steps, na.rm = TRUE))
        }

bydate <- Bydate(activity)
names(bydate)[2] <- "total.steps"
Hist(bydate)
### Calculate and report the mean and median of the total number of steps taken per day
orig_mean <- mean(bydate$total.steps)
orig_median <- median(bydate$total.steps)
```
**The orignal mean value with NAs**

```{r echo = TRUE}
orig_mean
```

**The orignal median value with NAs**

```{r echo = TRUE}
orig_median
```


### What is the average daily activity pattern? ###

```{r echo = TRUE}
byinterval <- summarize(group_by(activity, interval), 
                        mean(steps, na.rm = TRUE))
names(byinterval)[2] <- "avgsteps"
with(byinterval, plot(interval, avgsteps, 
                      type = "l",
                      ylab = "Average Steps Across All Days",
                      xlab = "5 Minutes Interval",
                      col  = "blue")
     )
```

**Contains the maximum number of steps?**  

```{r echo = TRUE}
byinterval[which(byinterval$avgsteps == max(byinterval$avgsteps)),]
```

### Input the missing values ###

**Count how many NAs in the original dataset**

```{r echo = TRUE}

all_na <- is.na(activity$steps)
sum(all_na)

```

**Using mean for that 5-minute interval to fill the NA**

```{r echo = TRUE}
new_act <- activity
obs <- length(all_na)

for (i in 1:obs){
       if(is.na(new_act$steps[i])){
                interval <- new_act[i,3]
                new_act[i,1] <- byinterval[which(byinterval$interval == interval),2]    
       }
       
}


```

**calculate the new_act dataset with NAs filled by mean of 5-minute interval**

```{r echo = TRUE}
bydate <- Bydate(new_act)
names(bydate)[2] <- "total.steps"
Hist(bydate)
new_mean <- mean(bydate$total.steps)
new_median <- median(bydate$total.steps)
```

**The new mean value with NAs filled**

```{r echo = TRUE}
new_mean
```

**The new median value with NAs filled**

```{r echo = TRUE}
new_median
```

###The findings after the NA filling###

**Compare the mean/median value with the original dataset, we may find the values are different.**
**The new dataset with NA filling, the mean is close to median**

###Are there differences in activity patterns between weekdays and weekends?###

```{r echo = TRUE}

weekend <- c("Saterday", "Sunday")
wkdays_act <- mutate(new_act, wkdays = weekdays(as.Date(date)))

for (i in 1:obs){
        if(wkdays_act$wkdays[i] %in% weekend){
            wkdays_act$wkdays[i] <- "weekend"
                
        }else{
            wkdays_act$wkdays[i] <- "weekday"
        }
}

byinterval <- summarize(group_by(wkdays_act, interval, wkdays), 
                        mean(steps, na.rm = TRUE))
names(byinterval)[3] <- "means"
results <- qplot(x = interval, y = means, 
                 data = byinterval,
                 geom = "line",
                 facets = wkdays~. ,
                 ylab = "Number of Steps"
                 )

```
**See below for the difference**
```{r echo = TRUE }
results

```

